\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2021-02-03 main.tex
compile latex
writing draft

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[top=2.2cm,left=2.4cm,right=2.4cm,bottom=3cm]{geometry}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}

@@

\title{\begin{center}
\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm]
\end{center}}
\date{\today}
\author{}

@@

%\section*{Extended Data Figures}

@@

{\setlength{\bibsep}{0\baselineskip} \small
\bibliographystyle{unsrt}
\bibliography{\string~/Dropbox/Bibliography/ET}}

>>>>

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[top=2.2cm,left=2.4cm,right=2.4cm,bottom=3cm]{geometry}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

@@

\title{\begin{center}
\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm]
\end{center}}
\date{}
\author{}

@@

\section{Introduction}

In the evolutionary dynamics literature it is common that the expected payoffs are
used in order to estimate individuals payoffs, and as a result, their fixation
probabilities.

The aim of this work is to understand the effect of the payoffs on the evolution
of the population. More specifically, this works evaluates the evolution for differently
calculated stochastic payoffs.

\section{Model}

The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
simultaneously and independently make a decision to Cooperate (C) or to defect
(D). The payoffs for both players depend on their actions and the actions of
their opponents. More specifically, the payoffs are given by:

\begin{equation}\label{eq:pd_normal_form}
    \begin{pmatrix}
        R & S  \\
        T & P
    \end{pmatrix}
\end{equation}

The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).
A special case of the PD is that of the donation game where the matrix~(\ref{eq:pd_normal_form})
is now written as:

A special case is the donation game,” where each player can cooperate by providing 
a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:

\begin{equation}\label{eq:donation_normal_form}
    \begin{pmatrix}
        b-c & -c  \\
        b & 0
    \end{pmatrix}
\end{equation}

where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
the benefit.

The game becomes more interesting once repetition is taken into account, and
now te history of play is taken into account when making decisions. The iterated
form of the game is called the Iterated Prisoner's Dilemma (IPD). There are other
two players repeated games such as the Snowdrift game, Harmony and the Stag Hunt
game. These games are also taken into account in the numerical simulations section.

The space of strategies is infinite in the IPD. In this work we assume that
the individuals of our population can only follow a reactive strategy.

Reactive strategy are a set of memory-one strategies that only take into account
the previous action of the opponent. An example of a reactive strategy is Tit
For Tat. Reactive strategies can be written explicitly as a vector \(\in
\R_{3}\). More specifically a reactive strategy \(s\) is given by \(s=(y, p,
q)\) where \(y\) is the probability that the strategy opens with a cooperation
and \(p, q\) are the probabilities that the strategy cooperates given that the
opponent cooperated and defected equivalently.

Two players with reactive strategies takes the form of a Markov chain
with four possible states CC, CD, DC, DD (the possible outcomes of each round).
Given two reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and 
\(s_{2}=(y_{2}, p_{2}, q_{2})\), the transition matrix \(M\) of this Markov chain
is given by:

\begin{equation}
    \input{static/matrix}
\end{equation}

The long run steady state probability
vector \(\mathbf{v}\), which is the solution to \(\mathbf{v} M = \mathbf{v}\), can be
combined with the payoff of matrix (\ref{eq:pd_normal_form}) to give the expected
payoffs for each player. More specifically, the payoffs for a reactive
strategy \(s_1\) against an opponent \(s_2\), is given by:

\begin{equation}
    \mathbf{v}(s_1, s_2) \cdot u
\end{equation}

The basic setup is a population of \(N\)
individuals, where \(N\) is even, and where mutations are sufficiently rare such
that at any point in time at most two different strategies are present in the
population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
respectively.

Each step of the evolutionary process consists of two stages, a game stage and
an updating stage. In the game stage, each player is randomly matched with some
other player in the population to interact in one instance of the repeated
prisoner's dilemma. In the updating stage, two players are randomly drawn from
the population, a `learner' and a `role model'. Given that the learner's payoff
in the last round is $u_L\!\in\! \mathcal{U}$ and that the role model's last
round's payoff $u_M\!\in\! \mathcal{U}$, we assume the learner adopts the role
model's strategy with probability

\begin{equation} \label{Eq:rho}
\rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
\end{equation}

The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.

% \begin{equation}\label{eq:transition_matrix}
%     \resizebox{.5\hsize}{!}{$\input{tex/m_matrix.tex}$}
% \end{equation}

% \begin{equation}\label{eq:press_dyson_utility}
%     u_q(p) = v \cdot (R, S, T, P).
% \end{equation}

@@

{\setlength{\bibsep}{0\baselineskip} \small
\bibliographystyle{unsrt}
\input{bibliography.bbl}
